---
title: "HW3"
author: "Maya Spaur"
date: "10/9/2019"
output: github_document
---

#Problem 1
Loading the data

```{r}
library(tidyverse)
library(p8105.datasets)
library(viridis)

data("instacart")



instacart %>%
  group_by(aisle) %>%
  summarize(n= n()) %>%
  mutate(aisle_rank = rank(desc(n))) %>%
  filter(aisle_rank <=3)

```

Describing the data

The `Instacart` dataset has `r nrow(instacart)` observations and 15 variables. The data is structured such that each row is a product from an order, and there is one order per user of the Instacart Online Grocery Shopping service. Key variables include order_id (order identifier), product_id (product identifier), product_name, and aisle. [What does it mean, give an illustrative example of observations --> give example of first item in table]?

There are `r n_distinct(pull(instacart, aisle_id))` unique types of aisles. The three most populare are fresh fruits, fresh vegetables, and packaged vegetables fruits. 



Plot with Number of Items per Aisle

```{r plot with number of items per aisle}
Number_of_items = 
  instacart %>%
  group_by(aisle) %>%
  summarize(n= n()) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = aisle, y = n)) + geom_col() + theme(axis.text.x = element_text(angle = 90)) + labs(x = "Aisle", y = "Number of Items Ordered", title = "Number of Items Ordered from Aisle")

Number_of_items

```

Table with 3 Most Popular Items

```{r table with 3 most popular items}
baking_ingredients =
instacart %>%
  filter(aisle== "baking ingredients") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
  

dog_food_care =
  instacart %>%
  filter(aisle == "dog food care") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)

packaged_vegetables_fruits =
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
         
binded_table = 
  rbind(baking_ingredients, dog_food_care, packaged_vegetables_fruits) %>%   knitr::kable()


binded_table
```

Time of Pink Lady Apples and Coffee Ice Cream Orders

```{r time of Pink Lady Apples and Coffee Ice Cream}
instacart %>%
  janitor::clean_names() %>%
  filter(product_name == "Pink Lady Apples"| product_name == "Coffee Ice Cream") %>%
  group_by(order_dow, product_name) %>%
  summarize(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>%
   knitr::kable(digits = 2)

```

COMMENTING ON THE RESULTS OF EACH ...

#Problem 2

Data Cleaning

```{r data cleaning}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health", response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(
    response = as.factor(response),
    response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
  

```

Answering questions 

```{r 2002 and 2007 states observed over 7 locations}

brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year == "2002") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
              filter(unique_location >= 7)
           
  
brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year == "2010") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
              filter(unique_location >= 7)
  
```

In 2002 there are 6 locations and in 2010 there are 14 locations.

Spaghetti Plot


```{r}
spaghetti_plot = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarize(
    avg_data_value = mean(data_value)
  ) %>%
  select(year, locationabbr, avg_data_value) %>%
  ggplot(aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) + geom_line() + theme(legend.position = "bottom") + labs(title = "Excellent Responses in US States", x = "Year", Y = "Avg Data Value")

spaghetti_plot

```

2 Panel Plot

```{r}
Two_panel_plot =
brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year== "2006" | year=="2010", locationabbr =="NY") %>%
  filter(topic == "Overall Health", response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
 ggplot(aes(x = response, y = data_value)) + geom_col() + facet_grid(~year)


Two_panel_plot
  
```


#Problem 3 How to make weekday and weekend variable?

Loading and tidying the data 

```{r problem 3 part 1}
accelerometer_data =
 read_csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
 pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    names_prefix ="activity_",
    values_to = "activity_counts"
  ) %>%
    mutate(
    weekend = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday")
    ) %>%
  mutate(
    minutes = as.integer(minutes)
  )

accelerometer_data

accelerometer_data2 =
accelerometer_data %>%
    group_by(day_id) %>%
  summarize(
    total_activity = sum(activity_counts))
  
accelerometer_data2

```

The `accelerometer_data`day dataset has [DESCRIBE THE RESULTING DATASEST], 1443 variables and 35 observations. Key variables include week and day.


Making a plot
#How to get this plot to work correctly
```{r}
accelerometer_data3 =
accelerometer_data %>%
  mutate(
    hour = minutes %/% 60,
    hour = as.integer(hour),
    unique_day = paste(week, day)
  ) %>%
  group_by(unique_day, hour) %>%
  summarize(mean_activity_counts = mean(activity_counts))

accelerometer_plot =
accelerometer_data3 %>%
  ggplot(aes(x = hour, y = mean_activity_counts, color = unique_day)) + geom_line()

accelerometer_plot

   
```

