---
title: "HW3"
author: "Maya Spaur"
date: "10/9/2019"
output: github_document
---

#Problem 1

Setup
```{r}
library(tidyverse)
library(p8105.datasets)
library(viridis)

theme_set(theme_bw() + theme(legend.position = "bottom"))

```

Loading the data 

```{r}
data("instacart")

instacart %>%
  group_by(aisle) %>%
  summarize(n= n()) %>%
  mutate(aisle_rank = rank(desc(n))) %>%
  filter(aisle_rank <=3)

```

Describing the data

The `Instacart` dataset has `r nrow(instacart)` observations and 15 variables. The data is structured such that each row is a product from an order, and there is one order per user of the Instacart Online Grocery Shopping service. Key variables include order_id (order identifier), product_id (product identifier), product_name, and aisle. For example, the first row of the table contains the product named Bulgarian Yogurt, which is located in the Yogurt Aisle in the Dairy Eggs department. This product is further identified by the order number, the day of the week (order_dow) and the hour of the day (order_hour_of_day) on which it was ordered, and the user id.

There are `r n_distinct(pull(instacart, aisle_id))` unique types of aisles. The three most populare are fresh fruits, fresh vegetables, and packaged vegetables fruits. 



Plot with Number of Items per Aisle

```{r plot with number of items per aisle}
Number_of_items = 
  instacart %>%
  group_by(aisle) %>%
  summarize(n= n()) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = aisle, y = n)) + geom_col() + theme(axis.text.x = element_text(angle = 90)) + labs(x = "Aisle", y = "Number of Items Ordered", title = "Number of Items Ordered from Aisle")

Number_of_items

```

Table with 3 Most Popular Items

```{r table with 3 most popular items}
baking_ingredients =
instacart %>%
  filter(aisle== "baking ingredients") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
  

dog_food_care =
  instacart %>%
  filter(aisle == "dog food care") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)

packaged_vegetables_fruits =
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
         
binded_table = 
  rbind(baking_ingredients, dog_food_care, packaged_vegetables_fruits) %>%   knitr::kable()


binded_table
```

The most popular items in the aisle for baking_ingredients were light brown sugar, pure baking soda, and cane sugar. For the dog food care aisle, the most popular items were snack sticks, organix chicken, and small dog biscuits. In the packaged vegetables and fruits aisle, the most popular items were organic baby spianch, organic raspberries, and organic blueberries.


Time of Pink Lady Apples and Coffee Ice Cream Orders

```{r time of Pink Lady Apples and Coffee Ice Cream}
instacart %>%
  janitor::clean_names() %>%
  filter(product_name == "Pink Lady Apples"| product_name == "Coffee Ice Cream") %>%
  group_by(order_dow, product_name) %>%
  summarize(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>%
   knitr::kable(digits = 2)

```

In comparing times of orders of Pink Lady Apples and Coffee Ice Cream, averaged across hours of the day, Pink Lady Apples had consistently fewer orders than Coffee Ice Cream for each day of the week.

#Problem 2

Data Cleaning

```{r data cleaning}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health", response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(
    response = as.factor(response),
    response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
  

```

Answering questions 

```{r 2002 and 2007 states observed over 7 locations}

brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year == "2002") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
              filter(unique_location >= 7)
           
  
brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year == "2010") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
              filter(unique_location >= 7)
  
  
```

In 2002, 6 states (CT, FL, MA, NC, NJ and PA) were observed at 7 or more locations. In 2010, 14 states (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA) were observed at over 7 or more locations.

Spaghetti Plot


```{r}
spaghetti_plot = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarize(
    avg_data_value = mean(data_value)
  ) %>%
  select(year, locationabbr, avg_data_value) %>%
  ggplot(aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) + geom_line() + labs(title = "Excellent Responses in US States", x = "Year", Y = "Avg Data Value") 

spaghetti_plot

```

2 Panel Plot

```{r}
Two_panel_plot =
brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(year== "2006" | year=="2010", locationabbr =="NY") %>%
  filter(topic == "Overall Health", response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
 ggplot(aes(x = response, y = data_value)) + geom_col() + facet_grid(~year) + labs(title = "Poor to Excellent Responses in NY State", x = "Response", Y = "Data Value")


Two_panel_plot
  
```


#Problem 3

Loading and tidying the data 

```{r problem 3 part 1}
accelerometer_data =
 read_csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
 pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    names_prefix ="activity_",
    values_to = "activity_counts"
  ) %>%
    mutate(
    weekend = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday")
    ) %>%
  mutate(
    minutes = as.integer(minutes)
  )

accelerometer_data


```

The `accelerometer_data` dataset has 50,400 observations and 6 variables. Data was collected on a male 63-year old patient disgnosed with congestive heart failure and admitted to the Advanced Cardiac Care Center of Columbia University Medical Center. Key variables include day_id (one day_id corresponding to each of the 35 days of data collection), day (day of the week), minutes, activity_counts (total activity for each minute of the day), and weekend (coded as weekend for Saturday and Sunday, and weekday for Monday - Friday).

Total Activity of the Day

```{r}

accelerometer_data2 =
accelerometer_data %>%
    group_by(day_id) %>%
  summarize(
    total_activity = sum(activity_counts))

  
accelerometer_data2 %>% knitr::kable()


```

Total activity was shown to vary over day_id. Levels seemed to stabilize for a few days at a time, then would drop or increase drastically.

Plots for Activity Counts

Mean Activity Count for each Unique Day
```{r mean activity counts over each unique day}
accelerometer_data3 =
accelerometer_data %>%
  mutate(
    hour = minutes %/% 60,
    hour = as.integer(hour),
    unique_day = paste(week, day)
  ) %>%
  group_by(unique_day, hour) %>%
  summarize(mean_activity_counts = mean(activity_counts))

accelerometer_plot =
accelerometer_data3 %>%
  ggplot(aes(x = hour, y = mean_activity_counts, color = unique_day)) + geom_line()

accelerometer_plot

```

Mean Activity County for each day of the week for all 35 days over 5 weeks demonstrate the variability in accelerometer activity counts. Broad trends include higher levels of activity from about hour 7 to hour 21, reflecting relatively normal waking hours.

```{r mean activity counts over each day of the week}

accelerometer_data4=
accelerometer_data %>%
  mutate(
    hour = minutes %/% 60,
    hour = as.integer(hour),
  ) %>%
  group_by(day, hour) %>%
  summarize(mean_activity_counts = mean(activity_counts))


accelerometer_plot2 =
accelerometer_data4 %>%
  ggplot(aes(x = hour, y = mean_activity_counts, color = day)) + geom_line()

accelerometer_plot2

   
```

Based on the plot above, mean activity counts for all days of the week tend to increase from hour 5 to hour 10 or 10 and remain relatively active during typical waking hours, until hour 20. Mean activity counts tended to be lowest from hour 22 to hour 5, indicating sleeping.

Mean Activity Count for Tuesday

```{r}
accelerometer_data5=
accelerometer_data %>%
  filter(day == "Tuesday") %>%
  mutate(
    hour = minutes %/% 60,
    hour = as.integer(hour),
    unique_day = paste(week, day)
  ) %>%
  group_by(unique_day, hour) %>%
  summarize(mean_activity_counts = mean(activity_counts))
  

accelerometer_plot3 =
accelerometer_data5 %>%
  ggplot(aes(x = hour, y = mean_activity_counts, color = unique_day)) + geom_line()

accelerometer_plot3

```

Based on the graph of mean activity counts per hour on Tuesday, average actvity increased steadily from approximately hour 5 to hour 7, remained relatively active from hour 8 to hour 20 (daytime hours, with an exception for Tuesday of week 3), and decreased to hour 24. 
