---
title: "HW3"
author: "Maya Spaur"
date: "10/9/2019"
output: github_document
---

#Problem 1
Loading the data

```{r}
library(tidyverse)
library(p8105.datasets)
library(dplyr)
library(ggridges)

data("instacart")

```

Describing the data

The `Instacart` dataset has `r nrow(instacart)` observations and 15 variables. The data is structured such that each row is a product from an order, and there is one order per user of the Instacart Online Grocery Shopping service. Key variables include order_id (order identifier), product_id (product identifier), product_name, and aisle. [What does it mean, give an illustrative example of observations]?

There are `r n_distinct(pull(instacart, aisle_id))` unique types of aisles. 


```{r plot with number of items per aisle}
instacart %>%
  group_by(aisle) %>%
  summarize(n= n()) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = n, y = aisle)) + geom_point(alpha = 0.5) + labs(title = "Number of Items per Aisle, Items > 10,000", x = "Number of Items Ordered", y = "Aisle Name")

```


```{r table with 3 most popular items}
baking_ingredients =
instacart %>%
  filter(aisle== "baking ingredients") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
  

dog_food_care =
  instacart %>%
  filter(aisle == "dog food care") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)

packaged_vegetables_fruits =
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  group_by(aisle, product_name) %>%
  summarize(n=n()) %>%
  arrange(desc(n)) %>%
  head(3)
         
rbind(baking_ingredients, dog_food_care, packaged_vegetables_fruits)

 
```

```{r}


```


#Problem 2

#how to include only one response from excellent to poor? which response..
```{r}
data("brfss_smart2010")

summary(brfss_smart2010)


brfss_smart2010 %>%
  janitor::clean_names() %>%
  count(topic)

brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  count(response)

```


